Practical Performance Evaluation of Entity Resolution Algorithms: Lessons Learned at PatentsView.org

### 2022-08-02 paper/main.tex

\renewcommand{\baselinestretch}{2}

@@

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{natbib}
\usepackage{amssymb, amsthm}
\usepackage{listings}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{authblk}
\usepackage{lineno}
\linenumbers

@@

\usepackage{xcolor}
\newcommand{\ob}[1]{{\color{purple}#1}}

@@

\begin{abstract}
This paper introduces a novel evaluation methodology for entity resolution algorithms. It is motivated by PatentsView.org, a {U.S. Patents and Trademarks Office patent data exploration tool} which disambiguates patent inventors using an entity resolution algorithm. We provide a data collection methodology and tailored performance estimators which account for sampling biases. Our approach is simple, practical and principled -- key characteristics which allow us to paint the first representative picture of PatentsView's disambiguation performance. This is used to inform PatentsView's users of the reliability of the data and to allow the comparison of competing disambiguation algorithms.
\end{abstract}

\section{Introduction}
Entity resolution (also called record linkage, deduplication, or disambiguation) is the task of identifying records in a database which refer to the same entity. An entity may be a person, a company, an object or an event. Records are assumed to contain partially identifying information about these entities. When there is no unique identifier (such as a social security number) available for all records, entity resolution becomes a complex problem which requires sophisticated algorithmic solutions \citep{Herzog2007, Christen2012, Dong2015, Ilyas2019, Christophides2019, Christen2019, Papadakis2021, Binette2022a}.

For instance, the U.S. Patents and Trademarks Office (USPTO) makes available patent data dating back to 1790 \ob{(numerized full-text data is available from 1976)}. However, there is no standard for uniquely identifying inventors on patent applications. The result is a set of ambiguous mentions to inventors, where a single person's name may be spelled in different ways on two applications, and where two different inventors with the same name may be difficult to distinguish. \ob{Inventor mobility further complicates the use of contextual information for disambiguation.} Much research has been done within the fields of economics, computer science and statistics to disambiguate inventor mentions \ob{in this situation} \citep{trajtenberg2008identification, ferreira2012brief, ventura2013methods, Li2014, Ventura2015, kim2016random, yang2017mixture, morrison2017disambiguation, muller2017semantic, traylor2017learning, balsmeier2018machine,tam2019optimal, monath2019scalable, doherr2021disambiguation}.

The Office of the Chief Economist at the USPTO and the American Institutes for Research (AIR) currently maintain \textbf{PatentsView.org}, a patents knowledge discovery service which provides disambiguated patents data based on statistical entity resolution \citep{toole2021patentsview}. Our paper is motivated by this disambiguation service. Specifically, we are interested in the problem of \textit{evaluating} the accuracy of PatentsView's disambiguation, for the purpose of informing users of the reliability of the data and in order to support methodological research to improve upon PatentsView's disambiguation algorithm. We expand on the evaluation problem below.

@@

The entity resolution evaluation problem is to extrapolate from observed performance on small samples to real performance in a database with millions of records. \cite{Wang2022} refers to this as bridging the ``reality-ideality gap" in entity resolution, where high performance on benchmark datasets often does not translate in the real world. Here, performance may be defined as any combination of commonly used evaluation metrics for entity resolution, such as precision and recall, cluster homogeneity and completeness, rand index, or generalized merge distance \citep{Maidasani2012}. These metrics can be computed on benchmark datasets for which we have a ground truth disambiguation. However, the key evaluation problem is to obtain estimates which are representative of performance on the full data for which no ground truth disambiguation is available. This is challenging for the following reasons.

First, entity resolution problems {do not scale linearly}. While it may be easy to disambiguate a small dataset, the opportunity for errors grows \textit{quadratically} in the number of records. As such, we may observe good performance of an algorithm on a small benchmark dataset, while the true performance of on the entire data may be something else entirely.\footnote{This particular effect of dataset size in entity resolution is explored in \cite{Draisbach2013} in the context of choosing similarity thresholds.} This is a problem that PatentsView.org currently faces. Despite encouraging performance evaluation metrics on benchmark datasets, with nearly perfect precision and recall reported in the latest methodological report \citep{Monath2021}, the data science team at AIR observes lower real-world accuracy. {This phenomenon is illustrated in example \ref{first_example} below.}

@@

Our approach, detailed in sections \ref{sec:approach} and \ref{sec:proposed_estimators}, addresses these challenges by putting forward novel cluster-based expressions for performance metrics which reflect various sampling schemes. Each of these representations immediately suggests simple estimators which properly account for the above issues.

\begin{example}[Bias of precision computed on benchmark datasets]\label{first_example}
    To exemplify the problem with the trivial use of performance evaluation metrics on benchmark datasets, we have carried out a simple experiment which is described in detail in appendix \ref{appendix:example}. In short, we considered evaluating a disambiguation algorithm by sampling ground truth clusters and computing pairwise precision\footnote{Pairwise precision is the probability that a pair of record predicted to be a match is indeed a match. See section \ref{sec:proposed_estimators} for a formal definition} on this set of sampled clusters. This is analogous to the way that many real-world benchmark data sets are obtained and typically used. In this experiment, we know that the disambiguation algorithm has a precision of $52\%$ for the entire dataset.

    \begin{figure}[h]
        \centering
        \includegraphics[width=5.5in]{first-example}
        \caption{Distribution of precision estimates versus the true precision of $52\%$ (shown as a dotted vertical line). Panel \textbf{A} shows the trivial precision estimates computed for sampled records. Panel \textbf{B} shows our proposed precision estimates which accounts for the sampling mechanism. Sample bias and root mean squared error (rmse) are reported on each figure.}
        \label{fig:precision_problem}
    \end{figure}

    In panel \textbf{A} of figure \ref{fig:precision_problem}, we see the distribution of precision estimates versus the true precision of $52\%$ shown as a dotted vertical line. Precision estimates are most of the time very close to $100\%$ and always higher than $80\%$, despite the truth being a precision of only $52\%$. In contrast, panel \textbf{B} shows the distribution of our proposed precision estimator which is nearly unbiased. Both precision estimators rely on exactly the same data. They only differ in how they account for the underlying sampling process and the extrapolation from small benchmark datasets to full data.

@@

The first is model selection and comparison. PatentsView.org continually works at improving disambiguation methodology. This requires choosing between alternative methods and evaluating the results of methodological experiments. Without sound evaluation methodology, decisions regarding the disambiguation algorithm may be misaligned from real-world use and real-world performance. Notably, for a performance metric such as the pairwise f-score, one algorithm may perform better than another on a small benchmark dataset, while the opposite may hold true for performance on the entire data. This problem happens on typical benchmark datasets obtained from randomly sampling blocks or randomly sampling clusters (see section \ref{sec:proposed_estimators} for a definition of different sampling mechanisms).

The second is adequate use of disambiguated data. PatentsView.org's disambiguation results have been used in numerous scientific studies \citep{toole2021patentsview}. These studies make assumptions about the reliability of the data which need to be validated and upheld. In short, users of disambiguated data need to understand its reliability in order to make scientifically appropriate use of it. Evaluation aims to provide this rigorous reliability information.

@@

Much of past work has focused on defining and using relevant clustering evaluation metrics. The topic of estimating performance from samples has received much less attention, usually focusing on importance sampling estimators based on record pairs. We review contributions to these two main topics in the paragraphs below.

@@

Pairwise precision and recall metrics were first reported in \cite{Newcombe1959}, with \cite{Bilenko2003, Christen2007} emphasizing the importance of precision-recall curves for algorithm evaluation. Now, there are issues with the use of pairwise precision and recall in entity resolution applications, such as the large relative importance of large clusters. As such, other clustering metrics have been proposed, including cluster precision and recall, cluster homogeneity and completeness, and generalized merge distances \citep{Michelson2009, Menestrina2010, Maidasani2012, Barnes2015}. Important practical issues regarding the use of aggregate metrics are discussed in \cite{Hand2018}.

{While our work focuses on estimating pairwise precision and recall, the general approach also applies to any cluster based performance evaluation metrics, such as those described in \cite{Michelson2009}.}

@@

Regarding the reliable estimation of performance metrics, \cite{Belin1995} first proposed a semi-supervised approach to calibrate error rates when using a Fellegi-Sunter model \citep{fellegi_theory_1969}. \cite{Marchant2017} proposed an adaptive importance sampling estimator to estimate precision and recall from sampled record pairs. Other approaches to the estimation of performance metrics are model-based, where estimated precision and recall can be obtained from predicted match probabilities between record pairs \citep{enamorado2019using}. However, these model-based approaches cannot be used when working with black-box machine learning models or ad hoc clustering algorithms.

{In contrast to the above, our approach to estimation is more practical than pairwise sampling and applies to any black-box disambiguation algorithms such as those used at PatentsView.}

@@

Our approach to estimating performance metrics is based on the use of benchmark datasets which are already existing or which can be collected in a cost-effective way. These datasets contain entity clusters corresponding to sampling records and recovering all associated instances, to directly sampling clusters, or to sampling blocks. For each of these sampling processes, we propose estimators which correct for the issues discussed in section \ref{sec:challenges}, are nearly unbiased, and easy to use in practice.

Our approach has the following advantages:
\begin{enumerate}
    \item It can leverage existing benchmark datasets as well as new datasets collected specifically for performance evaluation.
    \item It can easily be generalized to estimate other clustering metrics, such as cluster precision, cluster recall, cluster homogeneity and completeness, and other generalized merge distances.
    \item For evaluation, the review of entity clusters is much more efficient than the review of record pairs. We can achieve high accuracy with small samples without relying on any sophisticated sampling scheme.
\end{enumerate}

{Furthermore, our approach is novel. To our knowledge, we are the first to propose unbiased performance estimators based on cluster and block samples. Past work either ignored biases when computing precision and recall from benchmark datasets \citep{frisoli2018exploring, Monath2021, han2019disambiguating}, did not provide estimates for one of precision or recall \citep{mcveigh2019scaling}, or provided solutions that are tailored to very specific record linkage models \citep{Belin1995}. We provide the first general solution to entity resolution evaluation which does not rely on sampling record pairs and which applies to any disambiguation algorithm.}

In short, the proposed approach is simple, principled, and practical. It is simple to use, it is statistically principled in its account of sampling processes and in its uncertainty quantification, and it is practical in the way that it can provide cost-effective estimates for any disambiguation algorithm.

@@

The rest of the paper is organized as follows. In section \ref{sec:data_methods}, {we describe: motivating benchmark datasets, our data collection process for evaluation, the proposed estimators, and our simulation study used to validate the performance of our estimators.} Section \ref{sec:results} then shows the results: our performance estimates and results from the simulation study. Finally, section \ref{sec:discussion} summarizes the paper and discusses future research directions.

@@

In this section, {we introduce: motivating datasets used at PatentsView, our hand-disambiguation methodology, our proposed performance metric estimators, and our simulation framework used to compare estimators.} {Note that we focus on \textit{inventor} disambiguation throughout, rather than the related problems of assignee and location disambiguation.}

@@

\paragraph{Israeli Inventors Benchmark}
\cite{trajtenberg2008identification} disambiguated the U.S. patents of Israeli inventors which were granted between 1963 and 1999. A total of 6,023 Israeli inventors were identified for this time period with 15,310 associated patents.

@@

\subsection{Hand-Disambiguation Protocol}\label{sec:hand-disambiguation}

In addition to the above benchmark datasets, {we have carried out a data collection experiment. This was motivated by the evaluation of the current PatentsView inventor disambiguation using the estimators proposed in section \ref{sec:proposed_estimators}.}

@@

Two AIR staff were tasked with recovering inventors' patents given sampled inventor mentions. To do this, they were assisted by tools available on PatentsView.org. First, given a sampled inventor, the associated predicted cluster was reviewed \ob{and, if any, wrongly} assigned patents were removed. Next, PatentsView's search tools were used to find additional mentions of similarly named inventors. These were reviewed \ob{and inventor mentions were added} to the predicted cluster if appropriate. The two AIR staff had an initial training session, followed by a test run on 10 inventors, before carrying out the rest of the data collection. They worked independently so that two resulting datasets were obtained for the same inventor mentions. In section \ref{sec:results}, these are referred to as the \textbf{Staff 1} and \textbf{Staff 2} datasets.

Note that our data collection methodology is biased towards PantentsView's current disambiguation. {Indeed, we do not expect the staff to have found all errors or all missing inventor mentions from predicted clusters. The staff used their best judgment, supported by a thorough search, to resolve inventor mentions.} In cases where no errors were found, the current disambiguation was assumed to be correct. Performance estimates based on this data might therefore be slightly optimistic. {This should be acknowledged when reporting performance estimates to PatentsView.org users. Otherwise, for the purpose of improving the current disambiguation \ob{algorithm, this data} is still appropriate to use. It represents the most visible errors in the current disambiguation rather than the totality of them.}

@@

First, we define pairwise precision and recall in terms of the number of links between records. Let $\mathcal{D} = \{1, 2, 3, \dots, N\}$ index a set of records let $\mathcal{C}$ be the partition of $\mathcal{D}$ representing ground truth clustering, and let $\widehat{\mathcal{C}}$ be a set of predicted clusters. Now let $\mathcal{T}$ be the set of record pairs which appear in the same cluster in $\mathcal{C}$ (matching pairs), and let $\mathcal{P}$ be the set of record pairs which appear in the same predicted cluster in $\widehat{\mathcal{C}}$ (predicted links). Pairwise precision ($P$) and pairwise recall ($R$) are then defined as
\begin{equation}\label{eq:def_P_R}
    P = \frac{\lvert \mathcal{T} \cap \mathcal{P} \rvert}{\lvert \mathcal{P} \rvert}, \quad R = \frac{\lvert \mathcal{T} \cap \mathcal{P} \rvert}{\lvert \mathcal{T} \rvert}.
\end{equation}

@@

We now provide three alternative representations of precision and recall which correspond to the processes of sampling records, sampling true clusters, and sampling blocks. These will be used to obtain precision and recall estimators under these sampling processes.

@@

In the cluster sampling case, sampling probabilities are typically known only up to normalizing factors. This is due to the fact that the total number of true clusters and other aspects of the ground truth cluster distribution are unknown in practice. As such, we provide expressions for precision and recall which only require knowing sampling probabilities up to normalizing factors. This allows the consideration of sampling uniformly at random and sampling clusters with probability proportional to their size.

@@

\begin{remark}
    Lemma \ref{lemma:record_sampling_representation} and lemma \ref{lemma:cluster_sampling_representation} can be generalized to apply to any performance metric which can be expressed as a sum $\sum_{c \in \mathcal{C}} h(c, \widehat{\mathcal{C}})$, for some function $h$, or as a function of such sums. For instance, the use of the so-called \textit{cluster} precision and \textit{cluster} recall \citep{Barnes2015}, or of cluster homogeneity and completeness \citep{Barnes2015}, can be more appropriate in the presence of large clusters. We leave these generalizations as extensions of our work.
\end{remark}

@@

\begin{remark}
    Lemmas \ref{lemma:record_sampling_representation} -- \ref{lemma:block_sampling_representation} are formulated in terms of sampled ground truth clusters and sampled blocks which do not contain errors. However, given the duality between precision and recall (interchanging the roles between $\mathcal{C}$ and $\widehat{\mathcal{C}}$  interchanges precision and recall), the results also apply to sampling \textit{predicted} clusters.
\end{remark}

@@

All of the expressions for precision and recall in lemmas \ref{lemma:record_sampling_representation} -- \ref{lemma:block_sampling_representation} are either population means or ratios of population means. As such, they can be estimated using sample means and ratios of sample means. For readability, we present here a generic formula for an approximately unbiased estimator of the ratio of means, and then specify the needed quantities for each representation below. The estimator applies a first order bias correction to the ratio of sample means, based on a Taylor approximation. Approximate confidence intervals can be computed based on the variance estimator of the ratio of sample means by Taylor approximation, assuming the corrected estimator has small bias \citep{fuller2011sampling}. The generic formula for estimating the ratio of $T$-sized ``population'' (of records/clusters/blocks) means of the form
\begin{align}
    E=\frac{1/T\sum_{i=1}^{T} B_i}{1/T\sum_{i=1}^T A_i},
\end{align}
assuming we have sampled $n$ elements (records/clusters/blocks), is
\begin{align}\label{eq:bias_adjustment}
    \widehat{E} = \frac{\bar{B}_n}{\bar{A}_n} \left\{ 1 + \frac{\theta_{n,T}}{n(n-1)} \sum_{s=1}^n \frac{A_s}{\bar{A}_n} \left(\frac{B_s}{\bar{B}_n} - \frac{A_s}{\bar{A}_n} \right) \right\}, \; \bar{A}_n = \frac{1}{n} \sum_{s=1}^{n} A_s, \; \bar{B}_n = \frac{1}{n} \sum_{s=1}^{n} B_s
\end{align}
We note that an additional symbol $\theta_{n,T}$ is introduced for possible finite population correction when relatively large number of elements are sampled without replacement (see below). Classical adjustment is set to $\theta_{n,T} = (1 - \frac{n-1}{T-1})$ \citep{Cochran1977}. For practical purposes when $T$ is large, often $\theta_{n,T}=1$ will suffice. In fact, knowledge of $T$ is not needed at all as long as it is large enough relative to $n$, which is useful because the total number of true clusters/blocks is not known in advance. Confidence interval can be computed based on the variance estimate of the above, which is
\begin{align} \label{eq:estimated_variance}
    \widehat{V}(\widehat{E}) = \left( \frac{\bar{B}_n}{\bar{A}_n} \right)^2 \frac{\theta_{n,T}}{n(n-1)} \sum_{s=1}^n \left( \frac{A_s}{\bar{A}_n} - \frac{B_s}{\bar{B}_n}\right)^2.
\end{align}
The specific values for $A_s$, $B_s$, and $T$ in each of our representation are described in appendix \ref{sec:appendix}.

@@

Bias and root mean squared error are reported in table \ref{tab:precision_results}.

@@

>>>>

\renewcommand{\baselinestretch}{1.2}

@@

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{natbib}
\usepackage{amssymb, amsthm}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{authblk}
\usepackage{lineno}
\usepackage{url}
%\linenumbers

\usepackage{listings}
\usepackage{xcolor}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    commentstyle=\color{codegreen},
    keywordstyle=\bf,
    stringstyle=\color{magenta},
    basicstyle=\ttfamily,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=none,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

@@

\usepackage{xcolor}
\newcommand{\ob}[1]{{#1}}

@@

\begin{abstract}
This paper introduces a novel evaluation methodology for entity resolution algorithms. It is motivated by PatentsView.org, a {U.S. Patents and Trademarks Office patent data exploration tool} that disambiguates patent inventors using an entity resolution algorithm. We provide a data collection methodology and tailored performance estimators that account for sampling biases. Our approach is simple, practical and principled -- key characteristics that allow us to paint the first representative picture of PatentsView's disambiguation performance. This approach is used to inform PatentsView's users of the reliability of the data and to allow the comparison of competing disambiguation algorithms.
\end{abstract}

\section{Introduction}
Entity resolution (also called record linkage, deduplication, or disambiguation) is the task of identifying records in a database that refer to the same entity. An entity may be a person, a company, an object or an event. Records are assumed to contain partially identifying information about these entities. When there is no unique identifier (such as a social security number) available for all records, entity resolution becomes a complex problem which requires sophisticated algorithmic solutions \citep{Herzog2007, Christen2012, Dong2015, Ilyas2019, Christophides2019, Christen2019, Papadakis2021, Binette2022a}.

For instance, the U.S. Patents and Trademarks Office (USPTO) makes available patent data dating back to 1790 \ob{(digitized full-text data is available from 1976)}. However, there is no standard for uniquely identifying inventors on patent applications. The result is a set of ambiguous mentions of inventors, where a single person's name may be spelled in different ways on two applications and where two different inventors with the same name may be difficult to distinguish. \ob{Inventor mobility further complicates the use of contextual information for disambiguation.} Much research has been done within the fields of economics, computer science and statistics to disambiguate inventor mentions \ob{in this situation} \citep{trajtenberg2008identification, ferreira2012brief, ventura2013methods, Li2014, Ventura2015, kim2016random, yang2017mixture, morrison2017disambiguation, muller2017semantic, traylor2017learning, balsmeier2018machine,tam2019optimal, monath2019scalable, doherr2021disambiguation}.

The Office of the Chief Economist at the USPTO and the American Institutes for Research (AIR) currently maintain \textbf{PatentsView.org}, a patents knowledge discovery service that provides disambiguated patents data based on statistical entity resolution \citep{toole2021patentsview}. Our paper is motivated by this disambiguation service. Specifically, we are interested in the problem of \textit{evaluating} the accuracy of PatentsView's disambiguation, for the purpose of informing users of the reliability of the data and in order to support methodological research to improve upon PatentsView's disambiguation algorithm. We expand on the evaluation problem below.

@@

The entity resolution evaluation problem is to extrapolate from observed performance in small samples to real performance in a database with millions of records. \cite{Wang2022} refer to this as bridging the reality-ideality gap in entity resolution, where high performance on benchmark datasets often does not translate into the real world. Here, performance may be defined as any combination of commonly used evaluation metrics for entity resolution, such as precision and recall, cluster homogeneity and completeness, rand index, or generalized merge distance \citep{Maidasani2012}. These metrics can be computed on benchmark datasets for which we have a ground truth disambiguation. However, the key evaluation problem is to obtain estimates that are representative of performance on the full data, for which no ground truth disambiguation is available. This is challenging for the following reasons.

First, entity resolution problems {do not scale linearly}. While it may be easy to disambiguate a small dataset, the opportunity for errors grows \textit{quadratically} in the number of records. As such, we may observe good performance of an algorithm on a small benchmark dataset, while the true performance on the entire dataset may be something else entirely.\footnote{This particular effect of dataset size in entity resolution is explored in \cite{Draisbach2013} in the context of choosing similarity thresholds.} This is a problem that PatentsView.org currently faces. Despite encouraging performance evaluation metrics on benchmark datasets, with nearly perfect precision and recall reported in the latest methodological report \citep{Monath2021}, the data science team at AIR observes lower real-world accuracy. {This phenomenon is illustrated in example \ref{first_example} below.}

@@

Our approach, detailed in sections \ref{sec:approach} and \ref{sec:proposed_estimators}, addresses these challenges by putting forward novel cluster-based expressions for performance metrics that reflect various sampling schemes. Each of these representations immediately suggests simple estimators that properly account for the above issues.

\begin{example}[Bias of precision computed on benchmark datasets]\label{first_example}
    To exemplify the problem with the trivial use of performance evaluation metrics on benchmark datasets, we carried out a simple experiment that is described in detail in appendix \ref{appendix:example}. In short, we evaluated a disambiguation algorithm by sampling ground truth clusters and computing pairwise precision\footnote{Pairwise precision is the probability that a pair of record predicted to be a match is indeed a match. See section \ref{sec:proposed_estimators} for a formal definition} on this set of sampled clusters. This is analogous to the way that many real-world benchmark datasets are obtained and typically used. In this experiment, we know that the disambiguation algorithm has a precision of $52\%$ for the entire dataset.

    \begin{figure}[h]
        \centering
        \includegraphics[width=5.5in]{first-example}
        \caption{Distribution of precision estimates versus the true precision of $52\%$ (shown as a dotted vertical line). Panel \textbf{A} shows the trivial precision estimates computed for sampled records. Panel \textbf{B} shows our proposed precision estimates which accounts for the sampling mechanism. Sample bias and root mean squared error (rmse) are reported in each figure.}
        \label{fig:precision_problem}
    \end{figure}

    In panel \textbf{A} of figure \ref{fig:precision_problem}, we see the distribution of precision estimates versus the true precision of $52\%$ shown as a dotted vertical line. Precision estimates are usually very close to $100\%$ and always higher than $80\%$, despite the truth being a precision of only $52\%$. In contrast, panel \textbf{B} shows the distribution of our proposed precision estimator which is nearly unbiased. Both precision estimators rely on exactly the same data. They only differ in how they account for the underlying sampling process and the extrapolation from small benchmark datasets to the full data.

@@

The first is model selection and comparison. PatentsView.org continually works at improving disambiguation methodology. This requires choosing between alternative methods and evaluating the results of methodological experiments. Without sound evaluation methodology, decisions regarding the disambiguation algorithm may not align with real-world use and real-world performance. Notably, for a performance metric such as the pairwise f-score, one algorithm may perform better than another on a small benchmark dataset, while the opposite may hold true for performance on the entire data. This problem arises with typical benchmark datasets obtained from randomly sampling blocks or randomly sampling clusters (see section \ref{sec:proposed_estimators} for a definition of different sampling mechanisms).

The second is adequate use of disambiguated data. PatentsView.org's disambiguation results have been used in numerous scientific studies \citep{toole2021patentsview}. These studies make assumptions about the reliability of the data that need to be validated and upheld. In short, users of disambiguated data need to understand its reliability in order to make scientifically appropriate use of it. Evaluation aims to provide this rigorous reliability information.

@@

Much of the past literature has focused on defining and using relevant clustering evaluation metrics. The topic of estimating performance from samples has received much less attention, usually focusing on importance sampling estimators based on record pairs. We review the contributions to these two main topics below.

@@

Pairwise precision and recall metrics were first reported in \cite{Newcombe1959}, with \cite{Bilenko2003} and \cite{Christen2007} emphasizing the importance of precision-recall curves for algorithm evaluation. However, there are issues with the use of pairwise precision and recall in entity resolution applications, such as the large relative importance of large clusters. As such, other clustering metrics have been proposed, including cluster precision and recall, cluster homogeneity and completeness, and generalized merge distances \citep{Michelson2009, Menestrina2010, Maidasani2012, Barnes2015}. Important practical issues regarding the use of aggregate metrics are discussed in \cite{Hand2018}.

{While our work focuses on estimating pairwise precision and recall, the general approach also applies to any cluster-based performance evaluation metrics, such as those described in \cite{Michelson2009}.}

@@

Regarding the reliable estimation of performance metrics, \cite{Belin1995} first proposed a semi-supervised approach to calibrating error rates when using a Fellegi-Sunter model \citep{fellegi_theory_1969}. \cite{Marchant2017} proposed an adaptive importance sampling estimator to estimate precision and recall from sampled record pairs. Other approaches to the estimation of performance metrics are model based, where estimated precision and recall can be obtained from predicted match probabilities between record pairs \citep{enamorado2019using}. However, these model-based approaches cannot be used when working with black-box machine learning models or ad hoc clustering algorithms.

{In contrast, our approach to estimation is more practical than pairwise sampling and applies to any black-box disambiguation algorithms such as those used at PatentsView.}

@@

Our approach to estimating performance metrics is based on the use of benchmark datasets that already exist or that can be collected in a cost-effective way. These datasets contain entity clusters corresponding to either: (a) sampling records and recovering all associated instances, (b) directly sampling clusters, or (c) sampling blocks. For each of these sampling processes, we propose estimators that correct for the issues discussed in section \ref{sec:challenges}, are nearly unbiased, and are easy to use in practice.

Our approach has the following advantages:
\begin{enumerate}
    \item It can leverage existing benchmark datasets as well as new datasets collected specifically for performance evaluation.
    \item It can easily be generalized to estimate other clustering metrics, such as cluster precision, cluster recall, cluster homogeneity and completeness, and other generalized merge distances.
    \item For evaluation, the review of entity clusters is much more efficient than the review of record pairs. We can achieve high accuracy with small samples without relying on sophisticated sampling schemes.
\end{enumerate}

{Furthermore, our approach is novel. To our knowledge, we are the first to propose unbiased performance estimators based on cluster and block samples. Past work either ignored biases when computing precision and recall from benchmark datasets \citep{frisoli2018exploring, Monath2021, han2019disambiguating}, did not provide estimates for precision or recall \citep{mcveigh2019scaling}, or provided solutions tailored to very specific record linkage models \citep{Belin1995}. We provide the first general solution to entity resolution evaluation that does not rely on sampling record pairs and that applies to any disambiguation algorithm.}

In short, the proposed approach is simple, principled, and practical. It is simple to use, it is statistically principled in its account of sampling processes and uncertainty, and it is practical in the way that it can provide cost-effective estimates for any disambiguation algorithm.

@@

The rest of the paper is organized as follows. In section \ref{sec:data_methods}, {we describe benchmark datasets, our hand-disambiguation methodology for evaluation, the proposed estimators, and our simulation study that we use to validate the performance of our estimators.} Section \ref{sec:results} then presents our performance estimates and results from the simulation study. Section \ref{sec:discussion} summarizes the paper and explores future research directions.

@@

In this section, {we introduce the benchmark datasets used at PatentsView, our hand-disambiguation methodology, our proposed performance metric estimators, and the simulation framework that we use to compare estimators.} {Note that we focus on \textit{inventor} disambiguation throughout, rather than on the related problems of assignee and location disambiguation.}

@@

\paragraph{Israeli Inventors Benchmark}
\cite{trajtenberg2008identification} disambiguated the U.S. patents of Israeli inventors that were granted between 1963 and 1999. A total of 6,023 Israeli inventors were identified for this time period with 15,310 associated patents.

@@

\subsection{Hand-Disambiguation Methodology}\label{sec:hand-disambiguation}

In addition to considering the above benchmark datasets, {we have carried out hand-disambiguation of inventor mentions. This was motivated by the evaluation of the current PatentsView inventor disambiguation using the estimators proposed in section \ref{sec:proposed_estimators}.}

@@

Two AIR staff were tasked with recovering inventors' patents given sampled inventor mentions. First, given a sampled inventor, the associated predicted cluster was reviewed \ob{and any wrongly} assigned patents were removed. Next, PatentsView's search tools were used to find additional mentions of similarly named inventors. These inventor mentions were reviewed \ob{and added} to the predicted cluster, if appropriate. The two AIR staff had an initial training session, followed by a test run on 10 inventors, before carrying out the rest of the data collection. They worked independently, which resulted in two datasets being obtained for the same inventor mentions. In section \ref{sec:results}, these are referred to as the \textbf{Staff 1} and \textbf{Staff 2} datasets.

Note that our data collection methodology is biased toward PantentsView's current disambiguation. {Indeed, we did not expect the staff to have found all errors or all missing inventor mentions from the predicted clusters. The staff used their best judgment, supported by a thorough search, to resolve inventor mentions.} In cases where no errors were found, the current disambiguation was assumed to be correct. Performance estimates based on this data might therefore be slightly optimistic, which should be acknowledged when reporting performance estimates to PatentsView.org users. Otherwise, for the purpose of improving the current disambiguation \ob{algorithm, this data} is still appropriate to use. It represents the most visible errors in the current disambiguation rather than the totality of them.

@@

First, we define pairwise precision and recall in terms of the number of links between records. Let $\mathcal{D} = \{1, 2, 3, \dots, N\}$ index a set of records let $\mathcal{C}$ be the partition of $\mathcal{D}$ representing ground truth clustering, and let $\widehat{\mathcal{C}}$ be a set of predicted clusters. Now let $\mathcal{T}$ be the set of record pairs that appear in the same cluster in $\mathcal{C}$ (matching pairs), and let $\mathcal{P}$ be the set of record pairs that appear in the same predicted cluster in $\widehat{\mathcal{C}}$ (predicted links). Pairwise precision ($P$) and pairwise recall ($R$) are then defined as
\begin{equation}\label{eq:def_P_R}
    P = \frac{\lvert \mathcal{T} \cap \mathcal{P} \rvert}{\lvert \mathcal{P} \rvert}, \quad R = \frac{\lvert \mathcal{T} \cap \mathcal{P} \rvert}{\lvert \mathcal{T} \rvert}.
\end{equation}

@@

We now provide three alternative representations of precision and recall that correspond to the processes of sampling records, sampling true clusters, and sampling blocks. These representations will be used to obtain precision and recall estimators under these sampling processes.

@@

In the cluster sampling case, sampling probabilities are typically known only up to a normalizing factor. This is because the total number of true clusters and other aspects of the ground truth cluster distribution are unknown in practice. As such, we provide expressions for precision and recall that only require knowing the sampling probabilities up to a normalizing factor. This allows the consideration of sampling uniformly at random and sampling clusters with probability proportional to their size.

@@

\begin{remark}
    Lemma \ref{lemma:record_sampling_representation} and lemma \ref{lemma:cluster_sampling_representation} can be generalized to apply to any performance metric that can be expressed as a sum $\sum_{c \in \mathcal{C}} h(c, \widehat{\mathcal{C}})$, for some function $h$, or as a function of such sums. For instance, the use of the so-called \textit{cluster} precision and \textit{cluster} recall \citep{Barnes2015}, or of cluster homogeneity and completeness \citep{Barnes2015}, can be more appropriate in the presence of large clusters. We leave these generalizations as extensions of our work.
\end{remark}

@@

\begin{remark}
    Lemmas \ref{lemma:record_sampling_representation} -- \ref{lemma:block_sampling_representation} are formulated in terms of sampled ground truth clusters and sampled blocks that do not contain errors. However, given the duality between precision and recall (interchanging the roles between $\mathcal{C}$ and $\widehat{\mathcal{C}}$  interchanges precision and recall), the results also apply to sampling \textit{predicted} clusters.
\end{remark}

@@

All of the expressions for precision and recall in lemmas \ref{lemma:record_sampling_representation} -- \ref{lemma:block_sampling_representation} are either population means or ratios of population means. As such, they can be estimated using sample means and ratios of sample means. For readability, we present here a generic formula for an approximately unbiased estimator of the ratio of means and then specify the needed quantities for each representation below. The estimator applies a first order bias correction to the ratio of sample means, based on a Taylor approximation. Approximate confidence intervals can be computed based on the variance estimator of the ratio of sample means by Taylor approximation, assuming the corrected estimator has a small bias \citep{fuller2011sampling}. The generic formula for estimating the ratio of $T$-sized ``population'' (of records/clusters/blocks) means of the form
\begin{align}
    E=\frac{1/T\sum_{i=1}^{T} B_i}{1/T\sum_{i=1}^T A_i},
\end{align}
assuming we have sampled $n$ elements (records/clusters/blocks), is
\begin{align}\label{eq:bias_adjustment}
    \widehat{E} = \frac{\bar{B}_n}{\bar{A}_n} \left\{ 1 + \frac{\theta_{n,T}}{n(n-1)} \sum_{s=1}^n \frac{A_s}{\bar{A}_n} \left(\frac{B_s}{\bar{B}_n} - \frac{A_s}{\bar{A}_n} \right) \right\}, \; \bar{A}_n = \frac{1}{n} \sum_{s=1}^{n} A_s, \; \bar{B}_n = \frac{1}{n} \sum_{s=1}^{n} B_s
\end{align}
We note that an additional symbol $\theta_{n,T}$ is introduced for a possible finite population correction when relatively large number of elements are sampled without replacement (see below). Classical adjustment is set to $\theta_{n,T} = (1 - \frac{n-1}{T-1})$ \citep{Cochran1977}. For practical purposes when $T$ is large, $\theta_{n,T}=1$ will suffice. In fact, knowledge of $T$ is not needed at all as long as it is large enough relative to $n$, which is useful because the total number of true clusters/blocks is not known in advance. Confidence intervals can be computed based on the variance estimate of the above, which is
\begin{align} \label{eq:estimated_variance}
    \widehat{V}(\widehat{E}) = \left( \frac{\bar{B}_n}{\bar{A}_n} \right)^2 \frac{\theta_{n,T}}{n(n-1)} \sum_{s=1}^n \left( \frac{A_s}{\bar{A}_n} - \frac{B_s}{\bar{B}_n}\right)^2.
\end{align}
The specific values for $A_s$, $B_s$, and $T$ in each of our representation are described in appendix \ref{sec:appendix_estimators}.

@@

@@

\subsection{Estimators Implementation}\label{sec:implementation}
